Acoustic Scene Classification: \\ Supervised and Unsupervised Approaches using the Scattering Transform

### 2016-06-21 paper/structureScene.tex
mention fourth-order Gammatone
Anden -> Andén + Oxford comma in author list
write on stability

\author{Vincent Lostanlen, Gr\'egoire Lafay, Joakim Anden and Mathieu Lagrange}

@@

\subsection{Invariance and stability in acoustic scene classification}
The notion of invariance to translation plays an essential role in auditory scene classification.
Indeed, the starting time of the recordings are chosen arbitrarily, and thus do not convey any information about the class.
To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space prior to training the classifier.
From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ of  a signal $\boldsymbol{x}(t)$ --
a translation-invariant representation up to time skills of duration $T$ can be simply obtained
by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
set to $1/T$:
\begin{equation}
\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \phi) (t).
\end{equation}
As a downside, the transient information in $\boldsymbol{x_1}$ at finer scales than $T$ are lost by this low-pass filtering, hence a lack of discriminability in feature space.
To address this issue, the temporal scattering transform is to recover this information by convolving $x_1$ with wavelets whose center frequencies are above $1/T$, and subsequently applying complex modulus.
\subsection{Temporal scattering}
Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
dimensionless frequency $1$ and bandwidth $1/Q$.
A filter bank of wavelets is built by dilating $\boldsymbol{\psi}$
according to a geometric sequence of scales $2^{-\gamma_1/Q}$,
where the variable $\gamma$ takes integer values.
\begin{equation}
\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q} \boldsymbol{\psi}(-2^{\gamma_1/Q} t)
\end{equation}
We denote by $\boldsymbol{\psi_{\gamma_1}}(t)$ the resulting wavelets.

The wavelet transform $\boldsymbol{}$ of an audio signal
$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets:
$\boldsymbol{y_1}(t, \gamma_1) =
(\boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}})(t)$.
Applying pointwise complex modulus to $\boldsymbol{y_1}$ yields
the wavelet scalogram
$\boldsymbol{x_1}(t, \gamma_1) = \vert \boldsymbol{y_1}(t, \gamma_1)\vert$.
The wavelet scalogram bears resemblance with the constant-Q transform (CQT),
which is derived from the short-time Fourier transform by averaging the frequency
axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}$.
Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
theoretical limit of optimal time-frequency localization across the whole
frequency range, whereas the CQT has a fixed temporal resolution.
Therefore, the wavelet scalogram has a better temporal localization at high
frequencies, which allows to observe steady amplitude modulations.

>>>>

\author{Vincent Lostanlen, Gr\'egoire Lafay, Joakim And\'en, and Mathieu Lagrange}

@@

\subsection{Invariance and stability in acoustic scene classification}
The notion of invariance to translation plays an essential role in auditory scene classification.
Indeed, the starting time of the recordings are chosen arbitrarily, and thus do not convey any information about the class.
To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space prior to training the classifier.
From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ of  a signal $\boldsymbol{x}(t)$ --
a translation-invariant representation up to time skills of duration $T$ can be simply obtained
by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
set to $1/T$:
\begin{equation}
\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \phi) (t).
\end{equation}
As a downside, the transient information in $\boldsymbol{x_1}$ at finer scales than $T$ are lost by this low-pass filtering, hence a lack of discriminability in feature space.
To address this issue, the temporal scattering transform recovers this information by convolving $x_1$ with wavelets whose center frequencies are above $1/T$ and subsequently applying complex modulus.

By resorting to wavelet modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to small time warps,
in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2010}.
Besides invariance to translation, this stability property is crucial in signal classification, as it entails a guarantee of robustness.

\subsection{Temporal scattering}
Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
dimensionless frequency $1$ and bandwidth $1/Q$.
A filter bank of wavelets is built by dilating $\boldsymbol{\psi}$
according to a geometric sequence of scales $2^{-\gamma_1/Q}$,
where the variable $\gamma$ takes integer values.
\begin{equation}
\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q} \boldsymbol{\psi}(-2^{\gamma_1/Q} t)
\end{equation}
We denote by $\boldsymbol{\psi_{\gamma_1}}(t)$ the resulting wavelets.

The wavelet transform $\boldsymbol{}$ of an audio signal
$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets:
$\boldsymbol{y_1}(t, \gamma_1) =
(\boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}})(t)$.
Applying pointwise complex modulus to $\boldsymbol{y_1}$ yields
the wavelet scalogram
$\boldsymbol{x_1}(t, \gamma_1) = \vert \boldsymbol{y_1}(t, \gamma_1)\vert$.
The wavelet scalogram bears resemblance with the constant-Q transform (CQT),
which is derived from the short-time Fourier transform (STFT) by averaging the frequency
axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}$.
Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
theoretical limit of optimal time-frequency localization across the whole
frequency range, whereas the CQT has a fixed temporal resolution, constrained by the support of the STFT analyzing window.
Therefore, the wavelet scalogram has a better temporal localization at high
frequencies, which allows to observe steady amplitude modulations.

In all subsequent experiments,
$\boldsymbol{\psi}$ is designed as a fourth-order Gammatone
wavelet of quality factor $Q = 8$.
